Search Strategies for Functional Logic Programming

### 2012-01-18 papers/atps2012/paper.tex
atps update

\begin{figure}[ht]
\centering
\begin{tabular}{|@{~~}l@{~~}|*{7}{@{~~}r@{~~}|}}
\hline
Program  &   DFS  & IDS(+1)& IDS(*2)&  eDFS  &  eBFS  &eIDS(+1)&eIDS(*2)\\
\hline
PermSort &  13.12 & 841.58 &  32.77 &  38.48 &  66.82 &  72.64 &  46.79 \\
Last     &        &        &        &        &        &        &        \\
Half     &        &        &        &        &        &        &        \\
Graph    &   1.76 &  52.53 &   2.40 &   3.51 &   4.22 &   4.26 &   3.83 \\
HorseMan &   2.64 & 639.64 &   6.63 &   3.23 &   3.29 &   3.20 &   3.17 \\
MAC      &   6.80 & 544.05 &   7.84 &  21.19 &  18.34 &  18.59 &  21.03 \\
Queens   &  24.74 & 374.64 &  33.62 &  42.71 &  43.98 &  44.01 &  42.65 \\
\hline
\end{tabular}
\caption{Benchmarks: computing all values}
\label{fig:bench-all-solutions}
\end{figure}

Figure~\ref{fig:bench-all-solutions} shows the benchmark results
when all values of a given expression are computed.
All benchmark programs are non-deterministic programs.
``PermSort'' sorts a list containing 15 elements by enumerating
all permutations and selecting the sorted ones,
``Last'' computes the last element \code{x}
of a list \code{xs} containing 100,000 elements
by solving the equation \ccode{ys++[x] =:= xs}
(see Section~\ref{sec:curry}),
``Half'' computes the half \code{y} of a natural number \code{x}
(in Peano representation) by solving the equation \code{y+y=:=x},
``Embed'' searches for an embedding of a message
which is ``hidden'' in another message
(see \cite[Sect.~4.1]{Antoy10JSC}),
\ldots

@@

{\sc should we also show relative run times?}

As one can see, depth-first search is the most efficient
search strategy.
However, complete search strategies are viable alternatives.
Their overhead is acceptable taking into account the fact
that one need not to reason about the details of exploring the
search space. The behavior of iterative deepening is largely
influenced by the parameters.
A constant increment of the depth bound causes a big overhead
compared to doubling the depth bound in each iteration.
Breadth-first search seems also a good strategy
taking into account the size of memory provided by modern computers.
However, there are extreme cases when the search tree is wide
where iterative deepening with recomputation is the single
strategy which is able to compute a result.
This is shown in Figure~\ref{fig:bench-first-solution}
that contains benchmarks to compute only a first value
of an expression.

>>>>

\begin{figure}[ht]
\centering
\begin{tabular}{|@{~~}l@{~~}|*{7}{@{~~}r@{~~}|}}
\hline
Program  &   DFS  & IDS(+1)& IDS(*2)&  eDFS  &  eBFS  &eIDS(+1)&eIDS(*2)\\
\hline
PermSort &  13.12 & 841.58 &  32.77 &  38.48 &  66.82 &  72.64 &  46.79 \\
Last     &   0.10 & $>$10min &   0.26 &   0.34 &   0.18 &   6.58 &   0.38 \\
Half     &   0.67 & 561.45 &   1.62 &   1.43 &   1.22 &   1.86 &   1.71 \\
Graph    &   1.76 &  52.53 &   2.40 &   3.51 &   4.22 &   4.26 &   3.83 \\
HorseMan &   2.64 & 639.64 &   6.63 &   3.23 &   3.29 &   3.20 &   3.17 \\
MAC      &   6.80 & 544.05 &   7.84 &  21.19 &  18.34 &  18.59 &  21.03 \\
Queens   &  24.74 & 374.64 &  33.62 &  42.71 &  43.98 &  44.01 &  42.65 \\
\hline
\end{tabular}
\caption{Benchmarks: computing all values}
\label{fig:bench-all-solutions}
\end{figure}

Figure~\ref{fig:bench-all-solutions} shows the benchmark results
when all values of a given expression are computed
with various search strategies.
All benchmark programs are non-deterministic programs.
``PermSort'' sorts a list containing 15 elements by enumerating
all permutations and selecting the sorted ones,
``Last'' computes the last element \code{x}
of a list \code{xs} containing 10,000 elements
by solving the equation \ccode{ys++[x] =:= xs}
(see Section~\ref{sec:curry}),
``Half'' computes the half \code{y} of a natural number \code{x}
(in Peano representation) by solving the equation \code{y+y=:=x},
%``Embed'' searches for an embedding of a message
%which is ``hidden'' in another message
%(see \cite[Sect.~4.1]{Antoy10JSC}),
``Graph'' computes some path in a graph where the edges
are represented by a non-deterministic ``successor'' operation,
``HorseMan'' computes the numbers of horses and men
from given numbers of heads and feet by searching for
appropriate numbers in Peano representation,
``MAC'' solves the ``Missionaries and Cannibals'' puzzle
(see \cite[Sect.~3.1]{AntoyHanus02FLOPS})
for reasonable numbers of missionaries and cannibals
to obtain measurable run times,
and ``Queens'' computes safe placements of queens on a chess board.

@@

The table entries in Figure~\ref{fig:bench-all-solutions}
contain the run times in seconds to compute all values of the initial
expression.
As one can see, the top-level depth-first search is the most efficient
search strategy.
The encapsulated version of depth-first search (eDFS)
with the explicit construction of the search tree
causes some overhead, since the search strategy
is encoded in the source program rather than in the run-time system
as in the top-level search (DFS).
The benchmarks also show that
complete search strategies, like ``eBFS'' and ``eIDS(*2)''
are viable alternatives to the incomplete depth-first search strategy.
Their overhead (sometimes they are even faster than ``eDFS'',
but they are always slower than top-level search)
is acceptable taking into account the fact
that one need not to reason about the details of exploring the
search space. The behavior of iterative deepening is largely
influenced by the parameters.
A constant increment of the depth bound causes a big overhead
compared to doubling the depth bound in each iteration,
in particular, when the search tree is slim as in ``Last''.
This is even worse for top-level iterative deepening
which recomputes the initial expression in each iteration,
see ``IDS(+1)''.

Breadth-first search seems a good strategy
taking into account the size of memory provided by modern computers.
However, there are extreme cases when the search tree is wide
where iterative deepening with recomputation is the single
strategy which is able to compute a result.
This is shown in Figure~\ref{fig:bench-first-solution}
that contains benchmarks to compute only a first value
of an expression.
